---
title: "Taller 5"
author: "Santiago Aillón, Santiago Arévalo, Juan José Caballero y Laura Ortiz"
date: "2022-10-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1).

#### a. Determine the population principal components Y1 and Y2 for the covariance matrix.

```{r}
y1 <- c(5,2)
y2 <- c(2,2)

Sigma = cbind(y1,y2)

eigen_val_Sigma = eigen(Sigma)$values
eigen_vec_Sigma = eigen(Sigma)$vectors

eigen_vec_Sigma
eigen_val_Sigma
```

<br />

#### b. Also, calculate the proportion of the total population variance explained by the first principal component.

```{r}
prop_pc1_Sigma = eigen_val_Sigma[1]/(eigen_val_Sigma[1] + eigen_val_Sigma[2])
prop_pc1_Sigma
```


<br />


### 2). Convert the covariance matrix in Exercise 1 to a correlation matrix ρ.

#### a. Determine the principal components Y1 and Y2 from ρ and compute the proportion of total population variance explained by Y1.

```{r}
cor = cov2cor(Sigma)
cor

eigen_val_cor = eigen(cor)$values
eigen_vec_cor = eigen(cor)$vectors

eigen_vec_cor

prop_pc1_cor = eigen_val_cor[1]/(eigen_val_cor[1] + eigen_val_cor[2])
prop_pc1_cor
```

<br />

#### b. Compare the components calculated in Part a with those obtained in Exercise 1. Are they the same? Should they be?

#### bla bla bla

<br />

#### c. Compute the correlations ρY1,Z1, ρY1,Z2, and ρY2,Z1

```{r}
py1z1_cor = (eigen_vec_cor[1,1]*sqrt(eigen_val_cor[1]))/(sqrt(Sigma[1,1]))
py1z2_cor = (eigen_vec_cor[1,2]*sqrt(eigen_val_cor[1]))/(sqrt(Sigma[2,2]))
py2z1_cor = (eigen_vec_cor[2,1]*sqrt(eigen_val_cor[2]))/(sqrt(Sigma[1,1]))

correlations = data.frame(py1z1_cor, py1z2_cor, py2z1_cor)
correlations
```


<br />


### 3). Data on x1 = sales and x2 = profits for the 10 largest companies in the world were already listed.

```{r}
S = cbind(c(7476.45, 303.62), c(303.62, 26.19))
```

<br />

#### a. Determine the sample principal components and their variances for these data. (You may need the quadratic formula to solve for the eigenvalues of S.)

```{r}
eigen_val_S = eigen(S)$values
eigen_vec_S = eigen(S)$vectors

eigen_vec_S
```

<br />

##### POR TEOREMA LAS VAR SON:

```{r}
eigen_val_S
```

<br />

#### b. Find the proportion of the total sample variance explained by y1gorrito. 

```{r}
prop_y1_S = eigen_val_S[1]/(eigen_val_S[1] + eigen_val_S[2])
prop_y1_S
```

<br />

#### c. Compute the correlation coefficients r_y1gorrito,xk, k = 1, 2. What interpretation, if any, can you give to the first principal component?

```{r}
py1x1_S = (eigen_vec_S[1,1]*sqrt(eigen_val_S[1]))/(sqrt(S[1,1]))
py1x2_S = (eigen_vec_S[1,2]*sqrt(eigen_val_S[1]))/(sqrt(S[2,2]))

correlationsS = data.frame(py1x1_S, py1x2_S)
correlationsS
```


<br />


### 4). Convert the covariance matrix S in Exercise 3 to a sample correlation matrix R.

#### a. Find the sample principal components y1gorrito, y2gorrito and their variances.

```{r}
R = cov2cor(S)
R

eigen_val_R = eigen(R)$values
eigen_vec_R = eigen(R)$vectors

eigen_vec_R
eigen_val_R
```

<br />

#### b. Compute the proportion of the total sample variance explained by y1gorrito.

```{r}
prop_pc1_R = eigen_val_R[1]/(eigen_val_R[1] + eigen_val_R[2])
prop_pc1_R
```

<br />

#### c. Compute the correlation coefficients r_y1gorrito,zk, k = 1, 2. Interpret y1gorrito.

```{r}
py1x1_R = (eigen_vec_R[1,1]*sqrt(eigen_val_R[1]))/(sqrt(S[1,1]))
py1x2_R = (eigen_vec_R[1,2]*sqrt(eigen_val_R[1]))/(sqrt(S[2,2]))

correlationsR = data.frame(py1x1_R, py1x2_R)
correlationsR
```

<br />

#### d. Compare the components obtained in Part a with those obtained in Exercise 3(a) do you feel that it is better to determine principal components from the sample covariance matrix or sample correlation matrix? Explain.


<br />


### 5). The weekly rates of return for five stocks listed on the New York Stock Exchange are given in Table 1 (stock data.DAT).

```{r}
stock_data = read.table("/home/laura/Desktop/univ/univ2022-2/AEDmrda/talleres/T5/stock_data.DAT.txt")
summary(stock_data)
```

<br />

#### a. Construct the sample covariance matrix S, and find the sample principal components. (Note that the sample mean vector x is displayed in Example 8.5.)

```{r}
covData = cov(stock_data)

eigen_vec_Data = eigen(covData)$vectors
eigen_val_Data = eigen(covData)$values

eigen_vec_Data
```

<br />

#### b. Determine the proportion of the total sample variance explained by the first three principal components. Interpret these components.

```{r}
prop_pc1_Data = eigen_val_Data[1]/sum(eigen_val_Data)
prop_pc2_Data = eigen_val_Data[2]/sum(eigen_val_Data)
prop_pc3_Data = eigen_val_Data[3]/sum(eigen_val_Data)

proportions_Data = data.frame(prop_pc1_Data, prop_pc2_Data, prop_pc3_Data)
sum(proportions_Data)
```

#### c. Given the results in Parts a-b, do you feel that the stock rates-of-return data can be summarized in fewer than five dimensions? Explain.


<br />


### 6). Consider the census-tract data listed in Table 2 (census tract.DAT). Suppose the observations on X_5 = median value home were recorded in ten thousands, rather than hundred thousands, of dollars; that is, multiply all the numbers listed in the sixth column of the table by 10.

```{r}
census_track = read.table("/home/laura/Desktop/univ/univ2022-2/AEDmrda/talleres/T5/census_track.DAT.txt")
summary(census_track)
```

<br />

#### a. Construct the sample covariance matrix S for the census-tract data when X_5 = median value home is recorded in ten thousands of dollars.

```{r}
census_track$V5 = census_track$V5*10
summary(census_track)

cov_census = cov(census_track)
cov_census
```

<br />

#### b. Obtain the eigenvalue-eigenvector pairs and the first two sample principal components for the covariance matrix in Part a.

```{r}
eigen_vec_census = eigen(cov_census)$vectors
eigen_val_census = eigen(cov_census)$values

eigen_vec_census
eigen_val_census
```

<br />

#### c. Compute the proportion of total variance explained by the first two principal components obtained in Part b. Calculate the correlation coefficients, r_yi,xk, and interpret these components if possible.

```{r}
prop_pc1_census = eigen_val_census[1]/sum(eigen_val_census)
prop_pc2_census = eigen_val_census[2]/sum(eigen_val_census)

proportions_census = data.frame(prop_pc1_census ,prop_pc2_census)
sum(proportions_census)

prcomp(cov_census)
```


