---
title: "a"
author: "a"
date: "2022-10-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stats)
```


### 1. Determine the population principal components, $Y_{1}$, $Y_{2}$ for the covariance matrix. Also, calculate the proportion of the total population variance explained by the first principal component.

$$
\Sigma = \begin{bmatrix}
            5 & 2\\
            2 & 2
        \end{bmatrix}
$$

```{r}
col1 = rbind(5, 2)
col2 = rbind(2, 2)
# matriz de covarianzas y varianzas
sigma1 = cbind(col1, col2)
print(sigma1)
```

Valores y vectores propios:

```{r}
eigen_data_varmat = eigen(sigma1)
eigen_data_varmat
```

Porcion de la variabilidad de los datos explicada por el vector 1 de los vectores propios.

```{r}
eigen_data_var1 = eigen_data_varmat$values[1]/sum(eigen_data_varmat$values)
eigen_data_var1
```


### 2. Convert the covariance matrix in Exercise 1 to a correlation matrix ρ.

```{r}
correlation_mat1 = cov2cor(sigma1)
correlation_mat1
```

**A)  Determine the principal components $Y_1$ and $Y_2$ from $\rho$ and compute the proportion of total population variance explained by $Y_1$.**

Valores y vectores propios:
```{r}
eigen_data_cormat1 = eigen(correlation_mat1)
eigen_data_cormat1
```

Porcion de la variabilidad de los datos explicada por el vector 1 de los vectores propios.

```{r}
eigen_data_cor1 = eigen_data_cormat1$values[1]/sum(eigen_data_cormat1$values)
eigen_data_cor1
``` 


**B)  Compare the components calculated in Part a with those obtained in Exercise 1. Are they the same? Should they be?**

Hmmmmmm, no son los mismos y no deberían serlo porque la matriz de correlación me está hablando de un ángulo entre variables y con base a estos ángulos saco un vector, mientras que la matriz de de varianzas y covarianzas no lo está haciendo, además ambas matrices son diferentes, por lo que sus vectores propios y valores, no deberían ser los mismos.


**C) Compute the correlations, $\rho_{Y_{1}, Z_{1}}$, $\rho_{Y_{1}, Z_{2}}$ and $\rho_{Y_{2}, Z_{1}}$  **

Mirar pagina 225 del lb.


### 3. Data on $x_{1} =$ sales and $x_{2} =$ profits for the 10 largest companies in the world were already listed. We know that

$$ 
\bar{x} = \begin{bmatrix}
    155.60 \\
    14.70
    \end{bmatrix},  
S = \begin{bmatrix}
    7476.45 & 303.62 \\
    303.62 & 26.19
    \end{bmatrix} 
$$


```{r}
xbar = rbind(155.60, 14.70)
sigma2 = matrix(c(7476.45, 303.62, 303.62, 26.19), nrow=2, ncol=2, byrow = TRUE)
print(xbar)
print(sigma2)
```

**A) Determine the sample principal components and their variances for these data. (You may need the quadratic formula to solve for the eigenvalues of S). **

```{r}
eigen_data2 = eigen(sigma2)
eigen_data2
```

**B) Find the proportion of the total sample variance explained by $\hat{y}_{1}$.**

```{r}
variation_comp1 = eigen_data2$values[1]/sum(eigen_data2$values)
variation_comp2 = eigen_data2$values[2]/sum(eigen_data2$values)
print(variation_comp1)
print(variation_comp2)
```

**C) Compute the correlation coefficients $r_{\hat{y}_1,x_{k}}$, $k = 1, 2$. What interpretation, if any, can you give to the first principal component?**

AAAAAAAAAAAAAAAAAAAAAAAAAAa
AAAAAAAAAAAAAAAAAAAAAAAAAAa
AAAAAAAAAAAAAAAAAAAAAAAAAAa
AAAAAAAAAAAAAAAAAAAAAAAAAAa


### 4. Convert the covariance matrix S in Exercise 3 to a sample correlation matrix R.


```{r}
correlation_mat2 = cov2cor(sigma2)
correlation_mat2
```

**A) Find the sample principal components $\hat{y}_{1}$, $\hat{y}_{2}$ and their variances. **

```{r}
eigen_data_cormat2 = eigen(correlation_mat2)
eigen_data_cormat2
```

**B) Compute the proportion of the total sample variance explained by $\hat{y}_{1}$.**


```{r}
variation_comp1 = eigen_data_cormat2$values[1]/sum(eigen_data_cormat2$values)
variation_comp2 = eigen_data_cormat2$values[2]/sum(eigen_data_cormat2$values)
print(variation_comp1)
print(variation_comp2)
```


**C) Compute the correlation coefficients $r_{\hat{y}_1,x_{k}}$, $k = 1, 2$. What interpretation, if any, can you give to the first principal component?**

AAAAAAAAAAAAAAAAAAAAAAAAAAa
AAAAAAAAAAAAAAAAAAAAAAAAAAa
AAAAAAAAAAAAAAAAAAAAAAAAAAa
AAAAAAAAAAAAAAAAAAAAAAAAAAa


**D) Compare the components obtained in Part a with those obtained in Exercise 3 (a) do you feel that it is better to determine principal components from the sample covariance matrix or sample correlation matrix? Explain.**

Para estos datos particulares es mejor usar la matriz de covarianzas, pero puede que existan otros datos para los que sea mejor usar la matriz de correlación muestral.


### 5. The weekly rates of return for five stocks listed on the New York Stock Exchange are given in Table 1 ( stock data.DAT ).

```{r}
stock_data = read.delim("/home/turing/Documents/aed/stock_data.txt", sep = '\t')
stock_data
```

**B) Determine the proportion of the total sample variance explained by the first three principal components. Interpret these components. **

```{r}
sigma3 = cov(stock_data)
sigma3
```



```{r}
eigen_data3 = eigen(sigma3)
eigen_data3
```

```{r}
variation_comp1 = eigen_data3$values[1]/sum(eigen_data3$values)
variation_comp2 = eigen_data3$values[2]/sum(eigen_data3$values)
variation_comp3 = eigen_data3$values[3]/sum(eigen_data3$values)
print(variation_comp1)
print(variation_comp2)
print(variation_comp3)
```

La variabilidad capturada por todas estas componentes es:

```{r}
print(variation_comp1 + variation_comp2 + variation_comp3)
```

**C) Given the results in Parts a-b, do you feel that the stock rates-of-return data can be summarized in fewer than five dimensions? Explain. **

Si, podría resumirse en menos de 5 dimensiones, sin embargo 5 dimensiones ya es poco teniendo en cuenta que se quieren predecir datos como tasas de retorno del mercado mundial.


### 6. Consider the census-tract data listed in Table 2 (census tract.DAT). Suppose the observations on $X_{5} =$ median value home were recorded in ten thousands, rather than hundred thousands, of dollars; that is, multiply all the numbers listed in the sixth column of the table by 10.

**A) Construct the sample covariance matrix S for the census-tract data when $X5 =$ median value home is recorded in ten thousands of dollars. **


**B) Obtain the eigenvalue-eigenvector pairs and the first two sample principal components for the covariance matrix in Part a.**


**C) Compute the proportion of total variance explained by the first two principal components obtained in Part b. Calculate the correlation coefficients, ryi,xk, and interpret these components if possible**



