---
title: "Taller 7"
author: "Santiago Aillón, Santiago Arévalo, Juan José Caballero y Laura Ortiz"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Consider the two data sets
  
  $$
  X_1= \left[\begin{array}{cc} 
  3 & 7\\
  2 & 4\\
  4 & 7
  \end{array}\right]
  X_2= \left[\begin{array}{cc} 
  6 & 9\\
  5 & 7\\
  4 & 8
  \end{array}\right]
  $$
<br /> 

### for which


  $$
  \bar{x_1} = \left[\begin{array}{cc} 
  3\\
  6
  \end{array}\right]
  \bar{x_2} = \left[\begin{array}{cc} 
  5\\
  8
  \end{array}\right]
  $$  
<br /> 

### and


  $$
  S_{pooled} = \left[\begin{array}{cc} 
  1 & 1\\
  1 & 2
  \end{array}\right]
  $$ 
<br /> 

```{r}
X1 = cbind(c(3, 2, 4), c(7, 4, 7))
X2 = cbind(c(6, 5, 4), c(9, 7, 8))
X1

x1_bar = cbind(colMeans(X1))
x2_bar = cbind(colMeans(X2))

Spooled = cbind(c(1, 1), c(1, 2)) 
```

<br /> 

#### a). Calculate the linear discriminant function $\hat{y} = (\bar{x_{1}} − \bar{x_{2}})'S_{pooled}^{-1} x = \hat{a}'x$.

```{r}
y_gorrito = t(x1_bar - x2_bar)%*%solve(Spooled)
y_gorrito
```

$\hat{y} = (-2, 0) x_{0}$

<br /> 

#### b). Classify the observation $x_{0}'$= [ 2  7 ] as population $\pi_{1}$ or population $\pi_{2}$, using The Estimated Minimum ECM Rule for Two Normal Populations with equal priors and equal costs.

```{r}
m_gorrito = (1/2)*t(x1_bar - x2_bar)%*%solve(Spooled)%*%(x1_bar + x2_bar)
m_gorrito
```

```{r}
obs = c(2,7)
y0 = y_gorrito%*%obs
y0

if (y0 >= m_gorrito){
  sprintf("Como %.0f es mayor a %.0f entonces es de clase pi_1", y0, m_gorrito)

}else{
  sprintf("Como %.0f es menor a %.0f entonces es de clase pi_2", y0, m_gorrito)
}
```

<br /> 

#### c). Develop a linear classification function for the data using $\hat{y}$ from (a).


<br /> 

#### d). Using the function in (c) and $\hat{m} = \frac{1}{2} (\bar{x_{1}} − \bar{x_{2}})'S_{pooled}^{-1}(\bar{x_{1}} + \bar{x_{2}})$, construct the "confusion matrix" by classifying the given observations.

```{r}
n1C = 0 
n1M = 0
n2M = 0
n2C = 0
```

<br /> 

#### e). Given the results in (d), calculate the apparent error rate (APER).


<br /> 

### 2. Suppose that $n_{1} = 11$ and $n_{2} = 12$ observations are made on two random variables $X_{1}$ and $X_{2}$, where $X_{1}$ and $X_{2}$ are assumed to have a bivariate normal distribution with a common covariance matrix $\Sigma$, but possibly different mean vectors $\mu_{1}$ and $\mu_{2}$ for the two samples. The sample mean vectors and pooled covariance matrix are:

  $$
  \bar{x_1} = \left[\begin{array}{cc} 
  -1\\
  -1
  \end{array}\right]
  \bar{x_2} = \left[\begin{array}{cc} 
  2\\
  1
  \end{array}\right]
  $$  
  $$
  S_{pooled} = \left[\begin{array}{cc} 
  7.3 & -1.1\\
  -1.1 & 4.8
  \end{array}\right]
  $$ 
 
<br /> 

```{r}
x1_bar2 = c(-1,-1)
x2_bar2 = c(2,1)

Spooled2 = cbind(c(7.3, -1.1), c(-1.1,4.8))
```
 
#### a). Construct Fisher’s (sample) linear discriminant function: $\hat{y} = (\bar{x_{1}} − \bar{x_{2}})'S_{pooled}^{-1} x = \hat{a}'x$.

```{r}
y_gorrito2 = t(x1_bar2 - x2_bar2)%*%solve(Spooled2)
y_gorrito2
```

$\hat{y} = (-0.4906887, -0.5291162) x_{0}$

<br /> 

#### b). Assign the observation $x_{0}'$= [ 0  1 ] as population $\pi_{1}$ or population $\pi_{2}$. Assume equal costs and equal prior probabilities.

```{r}
m_gorrito2 = (1/2)*t(x1_bar2 - x2_bar2)%*%solve(Spooled2)%*%(x1_bar2 + x2_bar2)
m_gorrito2
```
```{r}
obs2 = c(0, 1)
y02 = y_gorrito2%*%obs2
y02

if (y02 >= m_gorrito2){
  sprintf("Como %.2f es mayor a %.2f entonces es de clase pi_1", y02, m_gorrito2)

}else{
  sprintf("Como %.2f es menor a %.2f entonces es de clase pi_2", y02, m_gorrito2)
}
```